\thetitle

### 2013-04-17 tfpie/tfpie.tex
more about profj

This also ultimately lead to the tool being abandoned
  as it was too difficult to maintain.)

The failures of  ProfJ:

\begin{itemize}
\item dead.
\item

\item commits to a particular fashion rather than enabling any.

\item abandons the first semester approach and replaces it with a
  (completely) object-oriented approach on day one.  There are
  analogies between functional and object concepts (``a method is like
  a function''), but the two never co-exist in a program.

\item Java, even pared down pedagogical Java, is syntactically heavy; discourages
  interactive programming

\end{itemize}

The changes that occur on Day 1 of HtDC w/ ProfJ:

\begin{itemize}
\item Radically different syntax
\item Types
\item Total commitment to objects
\item Subtly different base values
\end{itemize}

The changes that occur on Day 1 of HtDC w/o ProfJ:

\begin{itemize}
\item Different IDE
\item Compilation
\item Loss of interactivity
\end{itemize}

>>>>

\profj also takes on the the dual challenges of implementing Java as
well as subsetting it.  This ultimately resulted in both a limited
Java environment as well as the eventual abandoning of the tool since
it was too difficult to maintain, let alone keep up with advances in
Java.

Finally, committing to Java on the first day, regardless of the
environment provided to students, has significant limitations.  First,
the syntactic and semantic heaviness of Java is a burden for beginning
students, and discourages interactive and experimental programming.
The very first chapter of \htdc discusses the fixed size of Java integers,
a topic avoided entirely in the first course.
Second, by committing to a particular industrial-strength language, it
closes off possibilities in the curriculum.  Third, it commits
entirely to the new paradigm, making it more difficult for students to
compare the approaches.

Since \profj is no longer available, students are faced with an even
starker change on the first day.  Even with a student-oriented
environment such as DrJava or BlueJ~\cite{dvanhorn:Allen2002DrJava,
  dvanhorn:Hsia2005Taming}, students must learn an entirely new tool,
along with new libraries.
If the course uses a typical professional development environment such
as Eclipse, students must also contend with compilation, loss of
interactivity, and subtle issues such as classpaths, none of which are
fundamental to the concepts that the course focuses on.
